%% For tips on how to write a great abstract, have a look at
%%	-	https://www.cdc.gov/stdconference/2018/How-to-Write-an-Abstract_v4.pdf (presentation, start here)
%%	-	https://users.ece.cmu.edu/~koopman/essays/abstract.html
%%	-	https://search.proquest.com/docview/1417403858
%%  - 	https://www.sciencedirect.com/science/article/pii/S037837821830402X

\begin{abstract}
There has been extensive conversion of natural forest into farmland in southwest Western Australia since European settlement. This has led to significant changes in atmospheric patterns, and is implicated for an estimated 10\% loss of water vapour flows over this region. This work seeks to understand the mechanisms behind surface wind patterns surrounding these atmospheric changes. This work argues that beyond changes in surface friction, conversion of forest into farmland has fundamentally altered the thermodynamics of the region. We present evidence of how atmospheric energy conversions and surface energy fluxes can be sharply delineated by land cover differences. We present evidence suggesting a possible weakening of summer daytime sea breezes which resulted from historical clearing, which in theory would have led to decreased energy supply from coastal wind farms at the same time of increased energy demand from less cool relief for Perth. We also discover anomalies which point towards an unacknowledged contribution from atmospheric condensation. A set of general heuristics is formulated for assessing the likely interactions between future land cover change and circulations. A better understanding of these land-atmosphere interactions allows more accurate assessments of long-term project risk for wind energy developments, but also has import ramifications for cognate fields such as hydrology, climate science, environmental sustainability and hydroelectricity generation.

\paragraph{Author contribution} All code used was personally written from scratch, although small sections occasionally draw heavy inspiration from examples in online tutorials and help forums. All results and analysis were personally produced.
\end{abstract}