\begin{acknowledgements}
Thank you to my supervisor Dr Merlinde Kay for guiding me through my project and keeping me on task. 

Thank you to the UNSW IT support team, for answering my enquiries regarding use of the Katana computational cluster.

Thank you to the ECMWF specialist support team, for answering my enquiries regarding the ERA5 dataset.

Thank you to the NOAA-PSL data team, for answering my enquiries regarding their climate indices.

Thank you to the presenters who ran the American Meteorological Society's Python for Climate and Meteorology 2021 Short Course, including those from the Data Carpentry and MetPy teams, who provided me the basic understanding of relevant Python libraries to complete this thesis project. 

Thank you the development team for the xarray Python package (especially staff from the National Center for Atmospheric Research who created relevant tutorials), the Pangeo community, and the Project Pythia community, who created the various online resources and examples which I drew from in creating the code for this project. 

Thank you to the Australian Bureau of Meteorology, for providing access to the BARRA reanalysis dataset, and the NSW Department of Planning and Environment, for providing access to urban density datasets, even though a change in project scope meant these datasets were not used.

Thank you to Dr Jean-Paul Ebejer from The University of Malta, who created this \LaTeX thesis template and made it available for general use.

And in general, thank you to all the people who contributed to the literature, as well as software, which made this research possible.

This research includes computations using the computational cluster Katana supported by Research Technology Services at UNSW Sydney.
\end{acknowledgements}